\Huge{Complex State Management with Immutability}

### 2022-02-20 Paper.tex
Done Phase1

\subsection{I/O (especially networking and caching)}
%https://www.solarwinds.com/-/media/solarwinds/swdcv2/licensed-products/storage-resource-monitor/resources/whitepapers/top_4_causes_of_storage_io_bottlenecks.ashx
Applications that are I/O heavy often cause bottlenecks. As well as often causing the problem, I/O intensive applications are often more sensitive to a storage latency issue. When you have a large user base trying to access these applications, slowdowns tend to take place. Increased response time in storage I/O causes bottlenecks. When there is a queue in the storage I/O, you would generally see an increase in latency.
Network and Disk latency are often times the biggest source of headache if poorly managed, which is often times the case since they're usually treated as if they were function calls which they're clearly not.
% http://norvig.com/21-days.html#answers
\begin{center}
    \begin{table}[H]
        \begin{center}
        \begin{tabularx} {0.8\textwidth}{ 
            | >{\raggedright\arraybackslash}X 
        | >{\raggedleft\arraybackslash}X | }
        \hline
        execute typical instruction &	1/1,000,000,000 sec = 1 nanosec\\
\hline
    fetch from L1 cache memory &	0.5 nanosec\\
\hline
branch misprediction &	5 nanosec\\
\hline
    fetch from L2 cache memory &	7 nanosec\\
\hline
    Mutex lock/unlock &	25 nanosec\\
\hline
    fetch from main memory &	100 nanosec\\
\hline
    send 2K bytes over 1Gbps network &	20,000 nanosec\\
    \hline
    read 1MB sequentially from memory &	250,000 nanosec\\
\hline
    fetch from new disk location (seek) &	8,000,000 nanosec\\
\hline
    read 1MB sequentially from disk &	20,000,000 nanosec\\
\hline
    send packet US to Europe and back &	150 milliseconds = 150,000,000 nanosec\\
    \hline      
    % \float
\end{tabularx}

\end{center}
\caption{ \label{tab:1} Latency numbers every programmer should know}
\end{table}
\end{center}
As you can see disk fetches can take 100x more than main memory fetches. And when dealing with distributed systems were network call are an integral part of the system treating network call which 100000x more than main memory calls you end up with catastrophically unpredictable behavior

@@

Shadowing is a technique that could represent a changing variable. for instance, an accumulator. the technique is simple and possible in almost every programming language out there.
at it's simplest form shadowing looks like this:
\begin{lstlisting}
    // Scala  
    object Main {
        def main(args: Array[String]){
            val i = 1;
            {
                val i = 2;
                {
                    val i = 3;
                }
            }
        }  
    }    
\end{lstlisting}
But that's not really useful and even more confusing and very error prone and I'd agree shadowing in of it's self isn't really useful but it's really at the core of any recursive solution since every function come with it's own block and 
\begin{lstlisting} 
    def factorial (n: BigDecimal):BigDecimal = {
        if (n <= 1)  1
        else n * factorial(n-1)
    }
\end{lstlisting}
Notice here n value range over \{n, ... , 1\} but it's not really changing each n deffer from the other and has it's own scope
if you run this code it'll only go so far (around  n = 9613 for this example) until you get a StackOverflowError... not good.

@@

Another huge advantage lazy evaluation gives us

% \subsection{logfilesystem}
% sth
% \subsection{Append Logs}
% sth
% \subsection{Copy-On-Write}
% sth
% \subsection{Structural Sharing}
% effecient cow reduce the overhead of threading and parallel processing
% \subsection{MVC}
% sth
% \subsection{elm}
% sth
% \subsection{Cpp example}
% $https://www.youtube.com/watch?v=y_m0ce1rzRI\&t=4081s$
\section*{References}
[1]

@@

\section*{Acknowledgements}
The author wishes to thank Blabla and bla

>>>>

\subsection{I/O, Network and Disk}
%https://www.solarwinds.com/-/media/solarwinds/swdcv2/licensed-products/storage-resource-monitor/resources/whitepapers/top_4_causes_of_storage_io_bottlenecks.ashx
Applications that are I/O heavy often cause bottlenecks. As well as often causing the problem, I/O intensive applications are often more sensitive to a storage latency issue. When you have a large user base trying to access these applications, slowdowns tend to take place. Increased response time in storage I/O causes bottlenecks. When there is a queue in the storage I/O, you would generally see an increase in latency.
Network and Disk latency are often times the biggest source of headache if poorly managed, which is often times the case since they're usually treated as if they were function calls which they're clearly not.
% http://norvig.com/21-days.html#answers
\begin{center}
    \begin{table}[H]
        \begin{center}
        \begin{tabularx} {0.8\textwidth}{ 
            | >{\raggedright\arraybackslash}X 
        | >{\raggedleft\arraybackslash}X | }
        \hline
        execute typical instruction &	1/1,000,000,000 sec = 1 nanosec\\
\hline
    fetch from L1 cache memory &	0.5 nanosec\\
\hline
branch misprediction &	5 nanosec\\
\hline
    fetch from L2 cache memory &	7 nanosec\\
\hline
    Mutex lock/unlock &	25 nanosec\\
\hline
    fetch from main memory &	100 nanosec\\
\hline
    send 2K bytes over 1Gbps network &	20,000 nanosec\\
    \hline
    read 1MB sequentially from memory &	250,000 nanosec\\
\hline
    fetch from new disk location (seek) &	8,000,000 nanosec\\
\hline
    read 1MB sequentially from disk &	20,000,000 nanosec\\
\hline
    send packet US to Europe and back &	150 milliseconds = 150,000,000 nanosec\\
    \hline      
    % \float
\end{tabularx}

\end{center}
\caption{ \label{tab:1} Latency numbers every programmer should know[1]}
\end{table}
\end{center}
As you can see disk fetches can take 100x more than main memory fetches. And when dealing with distributed systems were network call are an integral part of the system treating network call which 100000x more than main memory calls you end up with catastrophically unpredictable behavior

@@

Shadowing is a technique that could represent a changing variable. for instance, an accumulator. the technique is simple and possible in almost every programming language out there.
at it's simplest form shadowing looks like this:
\begin{lstlisting}
    scala> object Main {
        def main(args: Array[String]) = {
            val i = 1;
            {
                val i = 2;
                {
                    val i = 3;
                }
            }
        }  
    }    
\end{lstlisting}
But that's not really useful and even more confusing and very error prone and I'd agree shadowing in of it's self isn't really useful but it's really at the core of any recursive solution since every function come with it's own block and 
\begin{lstlisting} 
    scala> def factorial (n: BigDecimal):BigDecimal = {
        if (n <= 1)  1
        else n * factorial(n-1)
    }
\end{lstlisting}
Notice here n value range over \{n, ... , 1\} but it's not really changing each n deffer from the other and has it's own scope
if you run this code it'll only go so far (around  n = 9613 for this example) until you get a StackOverflowError... not good.

@@

%###############################################
%######## Must Use Resource ####################
%###############################################
% Another huge advantage lazy evaluation gives us is the ability to use DAGs to optimize complex and large jobs
% https://techvidvan.com/tutorials/apache-spark-dag-directed-acyclic-graph/

% \subsection{Structural Sharing}
% \subsection{Cpp example}
% $https://www.youtube.com/watch?v=y_m0ce1rzRI\&t=4081s$
% \subsection{Copy-On-Write}
% \subsection{logfilesystem}
% \subsection{Append Logs}%not important
% effecient cow reduce the overhead of threading and parallel processing
% sth
% \subsection{MVC}
% \subsection{elm}
% sth
\section*{References}
[1] Peter Norvig $http://norvig.com/21-days.html\#answers$

@@

\section*{Acknowledgements}
Thank you for your time
